# Missingness in time: speaking the language of data {#ch:mists}

Can missing values methods speak the language of data, to go beyond the probabilistic taxonomy, MCAR, MAR, MNAR? Yes they can! The **naniar** package provides a suite of exploratory methods and imputation choices for tidy missing data analysis, that fit neatly into a tidy data workflow. However, temporal data is left behind. This talk will present exploratory methods for temporal missing, equipped with a new data-centric taxonomy. The work will be available in the R package called **mists**.

```{r mists-initial, echo = FALSE, cache = FALSE, include = FALSE}
read_chunk('R/mists.R')
```

```{r load-mists}
```

```{r count-na}
```

```{r miss-types-funs}
```

## Introduction

## Process for generating missing data in time

```{r miss-types, fig.height = 7}
```

## Tools for exploring temporal patterns of missingness

```{r miss-types-acf, fig.height = 6}
```

## Assessing results of imputation techniques

## Case study

## Discussions
