# Data representation, visual and analytical techniques for demystifying temporal missing data {#ch:mists}

Missing data provokes an air of mystery, that makes analysts itch throughout the exploration loop of transformation, visualization, and modeling. How to handle missing values involves decisions with many degrees of freedom, lending itself to a tedious and unwieldy process. The challenge of missingness roots in seeing what isn't there. The aim of this work is to clear that mysterious air away from missing data with the focus of temporal contexts from the data-centric perspective. A new sparse representation facilitates to index the runs of missings in time efficiently, with supporting operations and visual methods. This places missing data solely in the spotlight, speaking for themselves. When too many missings are scattered across variables and observations over time, missing data polishing strategies are populated and formulated. This equips analysts with tidy tools to iteratively remove missings from rows and columns, while keeping the temporal nature intact. The accompanying software is the R package **mists**.

```{r mists-initial, echo = FALSE, cache = FALSE, include = FALSE}
read_chunk('scripts/mists.R')
```

```{r load-mists}
```
