# Data representation, visual and analytical techniques for demystifying temporal missing data {#ch:mists}

Missing data provokes an air of mystery, that makes analysts itching throughout the exploration loop of transformation, visualization, and modelling. How to handle missing values involves decisions with many degrees of freedom, lending itself to a tedious and unwieldy process. The challenge of missingness roots in seeing what isn't there. The aim of this work is to clear that mysterious air away from missing data with the focus of temporal contexts from the data-centric perspective. A new sparse representation facilitates to index the runs of missings in time efficiently, with supporting operations and visual methods. This places missing data solely in the spotlight, speaking for themselves. When too many missings are scattered across variables and observations over time, missing data polishing strategies are populated and formulated. This equips analysts with tidy tools to iteratively remove missings from rows and columns, while keeping the temporal nature intact. The accompanying software is the R package **mists**

```{r mists-initial, echo = FALSE, cache = FALSE, include = FALSE}
read_chunk('scripts/mists.R')
```

```{r load-mists}
```

```{r count-na}
```

```{r miss-types-funs}
```

## Introduction

```{r miss-types, fig.height = 7, fig.cap = "(ref:miss-types-cap)"}
```

(ref:miss-types-cap) Time series lines along with missing distributions plotted for monthly number of international airline passengers from 1949 to 1960. This series shows an upward trend with strong seasonal patterns. The gray blocks indicate the missingness. To generate missing data in time, the process could be classified into four types of temporal missings: (1) missing at level, (2) missing at occasional, (3) missing at runs, (4) missing at season.
